---
title: "01_query_data"
author: "Markus Min"
date: "`r Sys.Date()`"
output: html_document
---

### Description
This Rmd file will go through how we queried PTAGIS and cleaned the resulting queries to prepare the data for later analyses.

## 1. Get tag codes for fish that meet our inclusion criteria

### 1.1 Initial Query

To determine which populations (natal origins) to include in our analysis, we first examined all adult returns (detections in the adult fishway at BON) since 2005.

This first query was created using the following steps:

From the [PTAGIS Dashboard](https://dashboard.ptagis.org): Advanced Reporting Home Page > Create Query Builder2 Report > Interrogation Summary

Once in Query Builder2, select the following filters:
- 1 Attributes: To the attributes that are selected by default (Tag, Mark File, Site, First Time, Last Time, First Antenna Group, Last Antenna Group, Release Date, Species, Run, Rear Type, and Release Site), and the following attributes: Release Site Basin, Release Site Subbasin, Mark Site Basin, Mark Site Subbasin, Release Site Longitude, Release Site Latitude, Mark Site, Hatchery, Stock, Mark Capture Method, Release Site Code, Mark Site Code, Mark Length, Mark Comment, Mark Conditional Comment, Mark Data Project, Mark Data Project Name, Mark Session Message
- 2 Metrics: Count, First Obs Date Max
- 16 First Obs Date/Time: 
  - Attribute: First Time
  - From: Value 
  - Is: Greater than or equal to
    - Value: 6/1/2005 12:00:00 AM
- 18 Obs Site: BO1 - Bonneville Bradford Is. Ladder, BO2 - Bonneville Cascades Is. Ladder, BO3 - Bonneville WA Shore Ladder/AFF, BO4 - Bonneville WA Ladder Slots, B2A - BONNEVILLE ADULT WA SHORE, BWL - Bonneville WA Shore Ladder
  - These are all of the Bonneville adult fishways and therefore will allow you to identify all adult returns
- 25 Species: Steelhead

Run Report, and download as a CSV. Save this in Data/PTAGIS_queries/intermediate_files/BON_sthd_2005-06-01_2025_01_23.csv
- The file is named to indicate the time period queried (from 2005-06-01 to the date the query was run)

### 1.2 Filter initial query

```{r chunk-options-libraries, message = FALSE, echo = FALSE, warning = FALSE}
# set chunk options
knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

```{r import_reformat_data}
adult_returns_complete <- clean_names(read.csv(here::here("Data", "PTAGIS_queries", "BON_sthd_2005-06-01_2025_01_23.csv")))

# Remove all individuals where the mark length is greater than 350 mm, but make sure to keep NAs
subset(adult_returns_complete, mark_length_mm <= 350 | is.na(mark_length_mm)) -> adult_returns

# Also remove all individuals with comments that indicate they were captured as adults
table(adult_returns$mark_capture_method_name)
as.data.frame(table(subset(adult_returns, is.na(mark_length_mm))$mark_conditional_comment_name)) %>% 
  arrange(desc(Freq))

# Keywords to identify individuals tagged as adults: mature, adult, returning

# Look at mark_session_message_value
as.data.frame(table(subset(adult_returns, is.na(mark_length_mm))$mark_session_message_value)) %>% 
  arrange(desc(Freq))
# Here you might also be able to ID based on words like "adult" and "reproduction" in the project description, but these are less reliable

# Find all mark conditional comments that contain the words "mature", "adult", or "returning"
unique_comments <- unique(adult_returns$mark_conditional_comment_name)
adult_comments <- unique_comments[grep("mature|Mature|adult|Adult|returning|Returning", unique_comments)]
# Take out the word immature
adult_comments <- adult_comments[!(adult_comments %in% c("Immature", "Caudal Fin Damage, Immature"))]

# Remove all fish with comments indicating that they were tagged as adults
adult_returns_complete %>% 
  subset(., mark_length_mm <= 350 | is.na(mark_length_mm)) %>%  # keep those with NA mark lengths or mark lengths <= 350 mm
  subset(., !(mark_conditional_comment_name %in% adult_comments)) -> adult_returns # remove individuals known to have been tagged as adults


nrow(subset(adult_returns, rear_type_code == "H" & is.na(mark_length_mm)))
nrow(subset(adult_returns, rear_type_code == "H" & !(is.na(mark_length_mm))))
# Approx 50% of hatchery hatchery individuals are missing a mark length


# Keep only certain columns
adult_returns %>% 
  distinct(tag_code, .keep_all = TRUE) %>% 
  dplyr::rename(mark_site_code = mark_site_code_value, release_site_code = release_site_code_value,
                mark_comment = mark_comment_name) %>% 
  # dplyr::select(tag_code, first_time_value, rear_type_code, release_site_name, release_site_basin_name, release_site_subbasin_name,
  #               mark_site_basin_name, mark_site_subbasin_name, mark_site_name, hatchery_name, stock_name, mark_site_code, release_site_code,
  #               mark_comment, mark_length_mm) %>% 
  mutate(time = mdy_hms(first_time_value)) %>% 
  mutate(date = date(time)) %>% 
  dplyr::select(-c(first_time_value, time))-> adult_returns

# Sort into run years
run_year <- c("05/06", "06/07", "07/08", "08/09", "09/10", "10/11", "11/12", "12/13", "13/14", "14/15", "15/16", "16/17", "17/18", "18/19", "19/20", "20/21", "21/22", "22/23", "23/24")
run_year_start <- seq(ymd("2005-06-01"), ymd("2023-06-01"), by = "years")
run_year_end <- seq(ymd("2006-05-31"), ymd("2024-05-31"), by = "years")

run_year_df <- data.frame(run_year, run_year_start, run_year_end)

# This code assigns fish to run years; you will receive this warning:
# Warning: Detected an unexpected many-to-many relationship between `x` and `y`.
# That's okay - what it's doing is duplicating every fish for every run year,
# but then the next step filters out all the incorrect run year assignments
adult_returns %>% 
  mutate(dummy =TRUE) %>% 
  left_join(run_year_df %>% mutate(dummy=TRUE), by = "dummy") %>% 
  filter(run_year_start <= date, run_year_end >= date) %>% 
  select(-c(dummy, run_year_start, run_year_end)) -> adult_returns

# Export this file as metadata
write.csv(adult_returns, here::here("Data", "covariate_data", "tag_code_metadata.csv"), row.names = FALSE)
```

### Selecting tributaries

As a starting point, we use the tributaries included in Richins & Skalski (2018):

1. Hood River
2. Fifteenmile Creek
3. Deschutes River
4. John Day River
5. Umatilla River
6. Walla Walla River
7. Yakima River
8. Wenatchee River
9. Entiat River
10. Tucannon River
11. Clearwater River
12. Grande Ronde River
13. Salmon River
14. Imnaha river

However, we will exclude the Hood River, as it is part of the Lower Columbia DPS.
Because of our emphasis on understanding complex movements, our focus
is on the Middle Columbia, Upper Columbia, and Snake River DPSs.

In addition to these tributaries, we will also include tributaries that have
at least 350 individuals across eight or more run years, that also have
sufficient arrays in the natal tributary to assess homing. Based on the below analysis,
we will add the Okanogan River, Methow River, and Asotin Creek populations. We will
not include the Klickitat and Wind River tributaries because we cannot 
reliably estimate detection efficiency in the tributary.

```{r}
# compare full dataset to Shelby's
as.data.frame(table(adult_returns$release_site_subbasin_name)) # full dataset
as.data.frame(table(adult_returns_subset$release_site_subbasin_name)) %>% 
  arrange(desc(Freq))

# New candidates for inclusion: Klickitat, Lochsa, Methow, Okanogan, Trout (or is this part of another river system?)

# Klickitat
klickitat_fish <- subset(adult_returns, release_site_subbasin_name == "Klickitat")
table(klickitat_fish$rear_type_code) 
table(subset(klickitat_fish, rear_type_code == "H")$run_year)
table(klickitat_fish$run_year)

# Lochsa
lochsa_fish <- subset(adult_returns, release_site_subbasin_name == "Lochsa")
table(lochsa_fish$rear_type_code) 
table(lochsa_fish$run_year)

# Methow
methow_fish <- subset(adult_returns, release_site_subbasin_name == "Methow")
table(methow_fish$rear_type_code) 
table(methow_fish$release_site_name)
table(methow_fish$run_year)
table(subset(methow_fish, rear_type_code == "W")$run_year)

# Okanogan
okanogan_fish <- subset(adult_returns, release_site_subbasin_name == "Okanogan")
table(okanogan_fish$rear_type_code) 
table(okanogan_fish$run_year)

# Trout
trout_fish <- subset(adult_returns, release_site_subbasin_name == "Trout")
table(trout_fish$rear_type_code) 
table(trout_fish$run_year)

deschutes_fish

# Upper columbia, Priest rapids
UCPA_fish <- subset(adult_returns, release_site_subbasin_name == "Upper Columbia-Priest Rapids")
as.data.frame(table(UCPA_fish$release_site_name))
# Ringold hatchery, right on the mainstem. Not a tributary

# Upper Salmon and Little salmon are part of the Salmon

# Middle Columbia - Hood
MCH_fish <- subset(adult_returns, release_site_subbasin_name == "Middle Columbia-Hood")
as.data.frame(table(MCH_fish$release_site_name))
# Mostly Hood, also Fifteenmile (already covered). But Wind River and Trout Creek (trib to Wind River) might be a pop?

wind_fish <- subset(adult_returns, release_site_name %in% c("WIND2R - Wind River, Washington", "TROUTC - Trout Creek (trib. to Wind River, Wash.)"))
table(wind_fish$rear_type_code) # 1070 wild
table(wind_fish$run_year)
table(wind_fish$release_site_name)

# Hells Canyon
HC_fish <- subset(adult_returns, release_site_subbasin_name == "Hells Canyon")
table(HC_fish$rear_type_code) 
table(HC_fish$run_year)
as.data.frame(table(HC_fish$release_site_name))

# Lower Snake
LS_fish <- subset(adult_returns, release_site_subbasin_name == "Lower Snake")
table(LS_fish$rear_type_code) 
table(LS_fish$run_year)
as.data.frame(table(LS_fish$release_site_name))
# No, all in mainstem of Snake

# Lower Snake - Asotin
LS_fish <- subset(adult_returns, release_site_subbasin_name == "Lower Snake-Asotin")
table(LS_fish$rear_type_code) 
table(LS_fish$run_year)
as.data.frame(table(LS_fish$release_site_name))
as.data.frame(table(subset(LS_fish, rear_type_code == "H")$release_site_name))
as.data.frame(table(subset(LS_fish, rear_type_code == "W")$release_site_name))

table(subset(LS_fish, rear_type_code == "W")$run_year)

# Chief Joseph
CJ_fish <- subset(adult_returns, release_site_subbasin_name == "Chief Joseph")
table(CJ_fish$release_site_name)
# All from Well's Hatchery, mainstem Columbia. 

# Similkameen
SK_fish <- subset(adult_returns, release_site_subbasin_name == "Similkameen")
table(SK_fish$rear_type_code)  # Just hatchery
table(SK_fish$run_year) # Not enough run years (only 8) - ignore

# Middle Columbia-Lake Wallula
wallula_fish <- subset(adult_returns, release_site_subbasin_name == "Middle Columbia-Lake Wallula")
table(wallula_fish$run_year)
table(wallula_fish$rear_type_code)
table(wallula_fish$release_site_name)
# Nope, mostly released in the mainstem
```


# PTAGIS - query all data

### Get tag codes for all fish of interest

The tributary systems that we want to query are:

1. Fifteenmile Creek
2. Deschutes River
3. John Day River
4. Umatilla River
5. Walla Walla River
6. Yakima River
7. Wenatchee River
8. Entiat River
9. Tucannon River
10. Clearwater River
11. Grande Ronde River
12. Salmon River
13. Imnaha river
14. Methow River
15. Okanogan River
16. Asotin Creek

#### Determine release site names for each tributary
```{r}
# Get name of all unique release sites
release_sites <- unique(adult_returns$release_site_name)

# Get release sites that correspond to each tributary

#  Fifteenmile Creek
fifteenmile_creek_sites <- release_sites[grep("Fifteen", release_sites)]

# Deschutes River
deschutes_river_sites <- unique(subset(adult_returns, release_site_basin_name == "Deschutes")$release_site_name)

# John Day River
john_day_river_sites <- unique(subset(adult_returns, release_site_subbasin_name %in% c("Upper John Day", "North Fork John Day", "Middle Fork John Day", "Lower John Day"))$release_site_name)

# Umatilla River
umatilla_river_sites <- unique(subset(adult_returns, release_site_subbasin_name == "Umatilla")$release_site_name)

# Walla Walla River
walla_walla_river_sites <- unique(subset(adult_returns, release_site_subbasin_name == "Walla Walla")$release_site_name)

# Yakima River
yakima_river_sites <- unique(subset(adult_returns, release_site_basin_name == "Yakima")$release_site_name)

# Wenatchee River
wenatchee_river_sites <- unique(subset(adult_returns, release_site_subbasin_name == "Wenatchee")$release_site_name)

# Entiat River
entiat_river_sites <- release_sites[grep("Entiat", release_sites)]

# Tucannon River
tucannon_river_sites <- release_sites[grep("Tucannon", release_sites)]

# Clearwater River
clearwater_river_sites <- unique(subset(adult_returns, release_site_basin_name == "Clearwater")$release_site_name)

# Grande Ronde River
grande_ronde_river_sites <- unique(subset(adult_returns, release_site_subbasin_name %in% c("Lower Grande Ronde", "Upper Grande Ronde", "Wallowa"))$release_site_name)

# Salmon River
salmon_river_sites <- unique(subset(adult_returns, release_site_basin_name == "Salmon")$release_site_name)

# Imnaha river
imnaha_river_sites <- unique(subset(adult_returns, release_site_subbasin_name == "Imnaha")$release_site_name)

# Methow River
methow_river_sites <- unique(subset(adult_returns, release_site_subbasin_name == "Methow")$release_site_name)

# Okanogan River
okanogan_river_sites <- unique(subset(adult_returns, release_site_subbasin_name == "Okanogan")$release_site_name)

# Asotin Creek
asotin_creek_sites <- release_sites[grep("Asotin", release_sites)]
```

#### Export table of release sites + natal origin

```{r}
data.frame(natal_origin = c(
  rep("Fifteenmile_Creek", length(fifteenmile_creek_sites)),
  rep("Deschutes_River", length(deschutes_river_sites)),
  rep("John_Day_River", length(john_day_river_sites)),
  rep("Umatilla_River", length(umatilla_river_sites)),
  rep("Walla_Walla_River", length(walla_walla_river_sites)),
  rep("Yakima_River", length(yakima_river_sites)),
  rep("Wenatchee_River", length(wenatchee_river_sites)),
  rep("Entiat_River", length(entiat_river_sites)),
  rep("Tucannon_River", length(tucannon_river_sites)),
  rep("Clearwater_River", length(clearwater_river_sites)),
  rep("Grande_Ronde_River", length(grande_ronde_river_sites)),
  rep("Salmon_River", length(salmon_river_sites)),
  rep("Imnaha_River", length(imnaha_river_sites)),
  rep("Methow_River", length(methow_river_sites)),
  rep("Okanogan_River", length(okanogan_river_sites)),
  rep("Asotin_Creek", length(asotin_creek_sites))
),
release_site_name = c(
                 fifteenmile_creek_sites, 
                 deschutes_river_sites,
                 john_day_river_sites,
                 umatilla_river_sites,
                 walla_walla_river_sites,
                 yakima_river_sites,
                 wenatchee_river_sites,
                 entiat_river_sites,
                 tucannon_river_sites,
                 clearwater_river_sites,
                 grande_ronde_river_sites,
                 salmon_river_sites,
                 imnaha_river_sites,
                 methow_river_sites,
                 okanogan_river_sites,
                 asotin_creek_sites
                 )) -> origin_table

# Export the origin table
write.csv(origin_table, here::here("Data", "covariate_data", "natal_origin_table.csv"), row.names = FALSE)
```



#### Get tag codes for all fish in river systems
Here we will break these tag codes into files containing 5,000 unique tags. This will keep the complete tag history files under 100 MB so that they can be uploaded to GitHub, and should make PTAGIS smoother.
```{r}
adult_returns %>% 
  subset(., release_site_name %in% c(fifteenmile_creek_sites, deschutes_river_sites, john_day_river_sites, umatilla_river_sites, walla_walla_river_sites, yakima_river_sites, wenatchee_river_sites, entiat_river_sites, tucannon_river_sites, clearwater_river_sites, grande_ronde_river_sites, salmon_river_sites, imnaha_river_sites, methow_river_sites, okanogan_river_sites, asotin_creek_sites)) -> adult_returns_for_analysis

# Write tag codes to multiple text files
as.data.frame(unique(adult_returns_for_analysis$tag_code)) -> unique_tags
  
    ntags <- length(unique_tags[,1])
    nfiles <- ceiling(ntags/5000)
    
    # Make an empty list to store the different files
    file_list <- list()
    
    # Split the tags into multiple files
    for (i in 1:nfiles){
      file_list[[i]] <- unique_tags[(1 + 5000 * (i - 1)):(5000 * i),1][!(is.na(unique_tags[(1 + 5000 * (i - 1)):(5000 * i),1]))]
    }
    
    # export the files
    for (i in 1:nfiles){
      write.table(file_list[[i]], file = here::here("Data", "PTAGIS_queries", "tag_codes", paste0("tag_codes_", i, ".txt")), col.names = FALSE, row.names = FALSE, quote = FALSE) 
    }
```

Next, these tag code files were used to run new queries on PTAGIS.

From the [PTAGIS Dashboard](https://dashboard.ptagis.org): Advanced Reporting Home Page > Create Query Builder2 Report > Complete Tag History

For each of these, the reports paste0("CTH_tag_codes_", 1:13, Date) were run. Those reports have the following settings:
Query Builder2:
- 1 Attributes: Tag, Event Type, Event Site, Event Date, Event Release Site, Event Release Date, (these first six are all standard, the rest you have to add manually),
Event Date Time, Antenna, Antenna Group, Antenna Group Configuration
- 2 Metrics: CTH Count
- 29 Tag Code: Import text file that you exported from R

Each file is exported as a CSV in Data/PTAGIS_queries/complete_tag_histories

# Covariate queries
To navigate to the appropriate page on [CBR DART](https://www.cbr.washington.edu/dart/query/basin_conditions_hist):

Homepage > Basin conditions > Project all years

For each of the following dams:

- BON-Bonneville
- IHR-Ice Harbor
- LWG-Lower Granite
- MCN-McNary
- PRD-Priest Rapids
- RIS-Rock Island
- RRH-Rocky Reach
- WEL-Wells

Select each of the following data:

- "Outflow KCFS, Daily Average"
- "Spill KCFS, Daily Average"
- "Temperature C, Daily Average"
- "Temperature C, Tailrace, Daily Average"

Download as CSV in Data/covariate_data/CBR_DART
- Store in a folder with the dam abbreviation
- Name files with paste0(dam, variable, query date)

##### Bonus code

```{r}
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_1.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 1) -> tag_codes_1
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_2.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 2) -> tag_codes_2
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_3.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 3) -> tag_codes_3
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_4.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 4) -> tag_codes_4
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_5.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 5) -> tag_codes_5
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_6.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 6) -> tag_codes_6
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_7.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 7) -> tag_codes_7
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_8.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 8) -> tag_codes_8
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_9.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 9) -> tag_codes_9
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_10.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 10) -> tag_codes_10
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_11.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 11) -> tag_codes_11
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_12.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 12) -> tag_codes_12
read.table(here::here("Data", "PTAGIS_queries", "tag_codes", "tag_codes_13.txt")) %>% 
  dplyr::rename(tag_code = V1) %>% 
  mutate(file = 13) -> tag_codes_13
tag_codes_1 %>% 
  bind_rows(., tag_codes_2) %>% 
  bind_rows(., tag_codes_3) %>% 
  bind_rows(., tag_codes_4) %>% 
  bind_rows(., tag_codes_5) %>% 
  bind_rows(., tag_codes_6) %>% 
  bind_rows(., tag_codes_7) %>% 
  bind_rows(., tag_codes_8) %>% 
  bind_rows(., tag_codes_9) %>% 
  bind_rows(., tag_codes_10) %>% 
  bind_rows(., tag_codes_11) %>% 
  bind_rows(., tag_codes_12) %>% 
  bind_rows(., tag_codes_13) -> tag_code_file_finder

write.table(tag_code_file_finder, here::here("Data", "PTAGIS_queries", "tag_codes", "tag_code_file_finder.txt"), row.names = FALSE)
```


